\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
I would first like to thank Maya Heat Transfer Technologies, specifically Chris Jackson and Christian Ruel, for sponsoring my Master's degree. I am also grateful for the funds received by the FQRNT and CRSNG as part of the Industrial Innovation Scholarship.

I would also like to express my gratitude towards my academic supervisor, Prof. Siva Nadarajah, for his patience and immense knowledge. His continued guidance has helped me stay on the right path throughout the last two years. 

I have also benefited greatly from the support and knowledge of my colleagues at Maya, namely Peyman Khayatzadeh, Roland Rivard, Kaveh Mohamed, Jérémie Bisson, Mohsen Karimian and especially Lee Betchen. My fellow student colleagues at McGill's Computational Aerodynamics Group have also been very helpful, and I would like to thank Jeremy Schembri, Yao Jiang, Phillip Zwanenburg, Farshad Navah and Aditya Kashi for their insight. From the same group, my friend Doug Shi-Dong has truly been a pleasure to work next to and I am grateful for all of our fruitful discussions. 

I am forever grateful to my girlfriend, Gabrielle Denis-Larocque, for her moral and emotional support. I would also like to express my sincere gratitude towards her parents, Josée Larocque and André Denis.

Last but not least, this work would never have been possible without my parents, Adel Belhaouane and Olga Weyhaeghe, and brother, Hakim Belhaouane. I owe them everything. 
